---
title: "Team 25 Project Proposal: [Insert Project Name]"
subtitle: "AAI 1001 Team Project 25"
author: 
  - name: "Balandra Dyanel Lancea Estillore, 2502025"
  - name: "Abburu Sai Tanvi, 2503638"
  - name: "Joash Tan Ee, 2501571"
  - name: "Soh Wei Hao, 2502369"
  - name: "Muhammad Izwan Bin Azman, 2503488"

date: "2026-02-26"
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: true
    self-contained: true
---

```{r}
library(readr)
library(dplyr)


population <- read_csv("data/population.csv", show_col_types = FALSE)
male <- read_csv("data/male-population.csv", show_col_types = FALSE)

cleaned_population <- population %>%
  left_join(male,
            by = c("Entity", "Code", "Year"),
            suffix = c("_Total", "_Male")) %>%
  rename(Population = `all years_Total`, Male_Population = `all years_Male`)

#Validations on joined table
#1. Missing Value
missing_before <- colSums(is.na(cleaned_population))
missing <- cleaned_population %>%
  filter(is.na(Code)) %>%
  distinct(Entity)%>%
  arrange(Entity)

cleaned_population <- cleaned_population %>%
  filter(!is.na(Code))

#Check missing values after filtering
missing_after <- colSums(is.na(cleaned_population))
comparison <- data.frame(
  Before = missing_before,
  After = missing_after
)
print(comparison)

#2. Duplicates
repeated <- cleaned_population %>%
  count(Entity, Code, Year) %>%
  filter(n > 1)

#3. Gender Population Calculation
cleaned_population <- cleaned_population %>%
  mutate(Female_Population = Population - Male_Population)

gender_mismatch <- cleaned_population %>%
  filter(abs(Population - (Male_Population + Female_Population)) > 1)

#4. Negative Values
negative <- cleaned_population %>%
  filter(Population < 0 |
           Male_Population < 0 |
           Female_Population < 0)

#5. Year Type
str(cleaned_population)
cleaned_population <- cleaned_population %>%
  mutate(Year = as.integer(Year))
str(cleaned_population)

#6 Year coverage check
range(cleaned_population$Year)
coverage_issues <- cleaned_population %>%
  group_by(Entity, Code) %>%
  summarise(
    min_year = min(Year),
    max_year = max(Year),
    total_years = n(),
    .groups = "drop"
  ) %>%
  filter(min_year != 1950 | max_year != 2023 | total_years != 74)

#7. Outliers
summary(cleaned_population$Population)

#Final Check
list(
  count_repeated = nrow(repeated),
  count_gender_mismatch = nrow(gender_mismatch),
  count_negative = nrow(negative),
  count_coverage_issues = nrow(coverage_issues)
)
colSums(is.na(cleaned_population))

```

```{r}
library(dplyr)
library(readr)

cleaned_population <- cleaned_population %>%
  arrange(Entity, Year) %>%
  group_by(Entity) %>%
  mutate(
    Growth = Population - lag(Population),
    Growth_Percentage = (Population - lag(Population)) / lag(Population) * 100
  ) %>%
  ungroup()

head(cleaned_population)
```



```{r}

library(countrycode)
library(dplyr)


# Add continent
cleaned_population <- cleaned_population %>%
  mutate(
    Continent = countrycode(sourcevar = Entity,
                            origin = "country.name",
                            destination = "continent")
  )

# Check countries still missing continent
continent_missing <- cleaned_population %>%
  filter(is.na(Continent)) %>%
  distinct(Entity) %>%
  arrange(Entity)

if(nrow(continent_missing) > 0){
  cat("Countries not matched to a continent:\n")
  print(continent_missing)
} else {
  cat("All countries matched to a continent.\n")
}

# Summarize: how many countries per continent (excluding NA)
continent_summary <- cleaned_population %>%
  filter(!is.na(Continent)) %>%
  distinct(Entity, Continent) %>%
  count(Continent) %>%
  arrange(desc(n))

# Preview first few rows
head(continent_summary)

write_csv(cleaned_population, "data/cleaned_population_final.csv")
```
## 1. Background
Show and explain the chosen visualization example, and the story to tell

The visualization created by Our World in Data presents a world population cartogram, where each small square represents 500,000 people in 2018. Instead of using the true geographic size of countries, the map scales each country according to its population size, allowing viewers to quickly see where the majority of the world’s population is located. In this representation, highly populated countries such as China and India appear significantly larger compared to countries with smaller populations. The visualization effectively communicates the distribution of the global population and allows comparisons between countries and regions. However, the original visualization primarily focuses on overall population counts and growth without providing deeper insights into demographic differences within populations.
To extend this visualization, we plan to integrate gender-based population data provided by the same website to analyse how male and female populations grow at different rates across countries over time. For example, countries such as Japan and Germany have aging populations where female populations may grow faster due to higher life expectancy, while countries such as Nigeria may experience rapid growth across both genders due to higher birth rates. By incorporating gender data, the visualization can reveal patterns such as gender imbalance, demographic aging, and differences in population growth trends across regions. This would provide a more comprehensive understanding of global population changes, allowing users to not only observe where populations are concentrated but also how gender dynamics influence population growth in different countries over time.

Strength and Weaknesses
One major strength of this visualization is that it effectively represents population proportions by resizing countries based on population instead of land area. This allows highly populated countries such as China and India to stand out clearly, helping viewers quickly understand where most of the world’s population is concentrated. Another strength is the use of a clear quantitative encoding where each square represents 500,000 people. This consistent unit allows the viewer to interpret the sizes of countries more meaningfully and understand how population differences are visually calculated. Additionally, the visualization presents a high level of information density by including population figures for individual countries while also providing regional totals, allowing users to compare population sizes at both the country and continental levels within a single graphic.

Despite its effectiveness, the visualization also has several limitations. One issue is visual clutter, particularly in densely populated regions such as Europe where many small countries and labels are placed very close together, making them difficult to read. Small countries are also harder to notice because they are represented by only a few squares, which can make them visually insignificant compared to larger nations. Another limitation is that the colour choices mainly serve to differentiate neighbouring countries rather than encode meaningful categories such as regions or population groups, which reduces the potential informational value of colour. Finally, because the visualization is static, users cannot zoom in, filter regions, or interact with the data, which limits deeper exploration and makes it harder to examine specific countries or densely packed areas.


**Chosen Visualisation example:** What is the specific friction point? 
**Story to tell:** xx
**strengths and weaknesses:** xx

## 2. Proposal
*Describe your solution and high-level architecture.*

Detail the application or system you intend to build. 

**Propose improvements :** xx
**Details on the data sources you plan to use.:** xx
**Technical Stack:** xx 


