---
title: "Team 25 Project Proposal: [Insert Project Name]"
subtitle: "AAI 1001 Team Project 25"
author: 
  - name: "Balandra Dyanel Lancea Estillore, 2502025"
  - name: "Abburu Sai Tanvi, 2503638"
  - name: "Joash Tan Ee, 2501571"
  - name: "Soh Wei Hao, 2502369"
  - name: "Muhammad Izwan Bin Azman, 2503488"

date: "2026-02-26"
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: true
    self-contained: true
---

```{r}
library(readr)
library(dplyr)


population <- read_csv("data/population.csv", show_col_types = FALSE)
male <- read_csv("data/male-population.csv", show_col_types = FALSE)

cleaned_population <- population %>%
  left_join(male,
            by = c("Entity", "Code", "Year"),
            suffix = c("_Total", "_Male")) %>%
  rename(Population = `all years_Total`, Male_Population = `all years_Male`)

#Validations on joined table
#1. Missing Value
missing_before <- colSums(is.na(cleaned_population))
missing <- cleaned_population %>%
  filter(is.na(Code)) %>%
  distinct(Entity)%>%
  arrange(Entity)

cleaned_population <- cleaned_population %>%
  filter(!is.na(Code))

#Check missing values after filtering
missing_after <- colSums(is.na(cleaned_population))
comparison <- data.frame(
  Before = missing_before,
  After = missing_after
)
print(comparison)

#2. Duplicates
repeated <- cleaned_population %>%
  count(Entity, Code, Year) %>%
  filter(n > 1)

#3. Gender Population Calculation
cleaned_population <- cleaned_population %>%
  mutate(Female_Population = Population - Male_Population)

gender_mismatch <- cleaned_population %>%
  filter(abs(Population - (Male_Population + Female_Population)) > 1)

#4. Negative Values
negative <- cleaned_population %>%
  filter(Population < 0 |
           Male_Population < 0 |
           Female_Population < 0)

#5. Year Type
str(cleaned_population)
cleaned_population <- cleaned_population %>%
  mutate(Year = as.integer(Year))
str(cleaned_population)

#6 Year coverage check
range(cleaned_population$Year)
coverage_issues <- cleaned_population %>%
  group_by(Entity, Code) %>%
  summarise(
    min_year = min(Year),
    max_year = max(Year),
    total_years = n(),
    .groups = "drop"
  ) %>%
  filter(min_year != 1950 | max_year != 2023 | total_years != 74)

#7. Outliers
summary(cleaned_population$Population)

#Final Check
list(
  count_repeated = nrow(repeated),
  count_gender_mismatch = nrow(gender_mismatch),
  count_negative = nrow(negative),
  count_coverage_issues = nrow(coverage_issues)
)
colSums(is.na(cleaned_population))

```

```{r}
library(dplyr)
library(readr)

cleaned_population <- cleaned_population %>%
  arrange(Entity, Year) %>%
  group_by(Entity) %>%
  mutate(
    Growth = Population - lag(Population),
    Growth_Percentage = (Population - lag(Population)) / lag(Population) * 100
  ) %>%
  ungroup()

head(cleaned_population)
```



```{r}

library(countrycode)
library(dplyr)


# Add continent
cleaned_population <- cleaned_population %>%
  mutate(
    Continent = countrycode(sourcevar = Entity,
                            origin = "country.name",
                            destination = "continent")
  )

# Check countries still missing continent
continent_missing <- cleaned_population %>%
  filter(is.na(Continent)) %>%
  distinct(Entity) %>%
  arrange(Entity)

if(nrow(continent_missing) > 0){
  cat("Countries not matched to a continent:\n")
  print(continent_missing)
} else {
  cat("All countries matched to a continent.\n")
}

# Summarize: how many countries per continent (excluding NA)
continent_summary <- cleaned_population %>%
  filter(!is.na(Continent)) %>%
  distinct(Entity, Continent) %>%
  count(Continent) %>%
  arrange(desc(n))

# Preview first few rows
head(continent_summary)

write_csv(cleaned_population, "data/cleaned_population_final.csv")
```
## 1. Background
*Define the context and the problem you are solving.*

In this section, provide a brief overview of the Fintech domain you are targeting (e.g., Blockchain, Personal Finance, or Regulatory Tech). 

* **Problem Statement:** What is the specific friction point? 
* **Motivation:** Why is this problem significant for the Singaporean or global market? 
* **Target Audience:** Who will benefit from this solution?

## 2. Proposal
*Describe your solution and high-level architecture.*

Detail the application or system you intend to build. 

* **Solution Overview:** How does your app solve the problem?
* **Core Functionalities:** List 3-4 key features.
* **Technical Stack:** Mention your use of Python, R, or Java. 

```{mermaid}
graph LR
    A[User Input] --> B{Fintech Engine}
    B --> C[Blockchain/API Layer]
    C --> D[Result/Visualization]
```



